---
layout: post
title: "Names of b.....s badder than Taylor Swift, a class in women's studies?"
comments: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE, 
                      cache = TRUE) 
```

[Once](http://www.masalmon.eu/2016/10/02/first7jobs-repost/) [again](http://www.masalmon.eu/2017/02/05/actuallivingscientists/), a Twitter trend sent me to my R prompt... Here is a [bit of context](http://knowyourmeme.com/memes/name-a-badder-bitch-than-taylor-swift). My summary: Taylor Swift apparently plays the bad girl in her new album and a fan of hers asked a question...

<blockquote class="twitter-tweet" data-lang="ca"><p lang="en" dir="ltr">Name a bitch badder than Taylor Swift 😍😛😤 <a href="https://t.co/AkSyQBUIME">pic.twitter.com/AkSyQBUIME</a></p>&mdash; Nutella (@xnulz) <a href="https://twitter.com/xnulz/status/928857792982781952?ref_src=twsrc%5Etfw">10 de novembre de 2017</a></blockquote>


The tweet was then quoted by many people mentioning badass women, and I decided to have a look at these heroes!

<!--more-->

# Name tweets badder than Nutella's

I was a bit lazy and asked Mike Kearney, `rtweet` maintainer, how to find tweets quoting a tweet, to which [Bob Rudis answered](https://twitter.com/hrbrmstr/status/937340032456814592). Now that I even had the code, it was no trouble at all getting the data. I added the filtering steps myself, see, I'm not _that_ lazy. I also removed the link to the quoted tweet that was at the end of each tweet. 

```r
question_tweet <- "928857792982781952"
badass <-  rtweet::search_tweets(question_tweet, n = 18000, include_rts = FALSE)
badass <- dplyr::filter(badass, is_quote_status, quote_status_id == question_tweet)
badass <- dplyr::mutate(badass, text = stringr::str_replace(text, "https://t\\.co/.*$", ""))
badass <- dplyr::mutate(badass, text = trimws(text))
readr::write_csv(badass, path = "data/2017-12-03-badderb_badass.csv")
```

```{r, echo = FALSE}
badass<- readr::read_csv("data/2017-12-03-badderb_badass.csv")
```

I obtained `r nrow(badass)` tweets. Not bad!

```{r}
library("magrittr")
set.seed(20171015)

indices <- sample.int(n = nrow(badass), size = 7)
badass$text[indices]

```

# Unnamed badder b.....s: your mother and grand-mother

Out of `r nrow(badass)`, `r sum(grepl("mother", badass$text))` contained the word "mother" -- I haven't looked for the word "mum" and haven't checked for the fact that it is someone from the family of the tweet author. Here a few of the personal stories (or not) identified with this quick and dirty method.

```{r}

set.seed(20171015)
mothers <- dplyr::filter(badass, stringr::str_detect(text, "mother"))
indices <- sample.int(n = 15)
mothers$text[indices]

```

Can we talk about that belly button thing?! I'm also happy to see a diversity of things they were recognized for. 


# Names of the badder b.....s

Quite a few of the tweets from this trend contained the name of someone. In order to extract these names, I resorted to a language processing method called entity extraction, the entity here being a person. For that, I could have used an [extractor module](https://app.monkeylearn.com/main/extractors/ex_isnnZRbS/) of the Monkeylearn platform [via my own `monkeylearn` package](https://github.com/ropensci/monkeylearn).

Instead, I chose to illustrate a different method: using the [`cleanNLP` package](https://github.com/statsmaths/cleanNLP) that I know from [the excellent R Journal paper presenting it](https://journal.r-project.org/archive/2017/RJ-2017-035/RJ-2017-035.pdf). Among other things, it serves as an interface between R and the [Python library spaCy](https://spacy.io/) and also as an interface between R and [the coreNLP Java library](https://stanfordnlp.github.io/CoreNLP/). Installing these tools is the painful part of the setup, but 1) you only need to install one of them 2) there are detailed instructions [here](https://github.com/statsmaths/cleanNLP#backends) 3) once your tool is installed, using the package is a breeze (and well independent of any rate limit contrary to `monkeylearn` use). I am at that breeze stage, you can be jealous. 
 
There were a few tweets with infuriating encoding issues, BOM or something like that, and I decided to just ignore them by using `purrr::possibly`. I obviously did this to illustrate the use of this `purrr` function, not out of laziness.
 
```r
library("cleanNLP")
init_spaCy()
# we need to remove characters like "\u0098"
badass <- dplyr::mutate(badass, text = enc2native(text))

get_entities_with_text <- function(x){
  obj <- run_annotators(x, as_strings = TRUE)
  entities <- get_entity(obj)
  entities$text <- x
  entities
}

possibly_get_entities <- purrr::possibly(get_entities_with_text,
                                         otherwise = NULL)

entities <- purrr::map_df(badass$text, possibly_get_entities)

readr::write_csv(entities, path = "data/2017-12-03-badderb_entities.csv")

```

```{r, echo = FALSE}
entities <- readr::read_csv("data/2017-12-03-badderb_entities.csv")
```

I got at least one entity for `r length(unique(entities$text))` out of `r nrow(badass)` tweets, and at least one _person_ for `r length(unique(entities[entities$entity_type == "PERSON",]$text))`. I am very satisfied with this.


# So, who are you, badder b.....s?

We get this kind of entities: `r unique(entities$entity_type)`. I'm more interested in PERSON and no, I'm not shouting. I chose to look at the top 12 in order to get a top 10 excluding Taylor Swift herself.

```{r}
entities %>%
  dplyr::filter(entity_type == "PERSON") %>%
  dplyr::group_by(entity) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::arrange(- n) %>%
  head(n = 12) %>%
  knitr::kable()

```

At that point I did feel like bursting out laughing though. Dora! And I checked, we're talking about Dora the explorer! Joan is Joan of arc. Interestingly in that top 10 we're mixing really bad persons, e.g. Myra Hindley was a serial killer, and really badass persons, like Rosa Parks. My husband will be happy to see Marie Curie in this list, since he's a big fan of hers, having even guided a few tours about her life in Paris.

Looking at the most frequently mentioned women obviously makes us loose well wrongly written names, and most importantly personal stories of badass mothers and the like, and of native women for instance, although I have the impression of having read about a few but probably because of my following [Auriel Fournier](https://twitter.com/RallidaeRule).

# Writing history?

I saw someone said they'd use the tweets as basis for history lessons. In order to get a view of a person, one could concatenate the tweets about them. Take Marie Curie for instance.

```{r}
entities %>%
  dplyr::filter(entity_type == "PERSON", entity == "Marie Curie") %>%
  dplyr::summarise(text = toString(text)) %>%
  .$text
```

Doing this one also gets the name of many other women. Moreover, if writing history lessons, one should have several sources, right? What about Wikidata [like in this other blog post of mine](http://www.masalmon.eu/2017/08/06/p1/)? It should have data for at least the most famous badass women.

```{r}
# add a function for getting a silent answer
quietly_query <- purrr::quietly(WikidataQueryServiceR::query_wikidata)

# function for getting someone's data
get_wikidata <- function(name, pb = NULL){
  if (!is.null(pb)) pb$tick()$print()
  Sys.sleep(1)
  item <- WikidataR::find_item(name, language = "en")
  # sometimes people have no Wikidata entry so I need this condition
  if(length(item) > 0){
    entity_code <- item[[1]]$id
    query <-  paste0("PREFIX entity: <http://www.wikidata.org/entity/>
                     #partial results
                     
                     SELECT ?propUrl ?propLabel ?valUrl ?valLabel ?picture
                     WHERE
                     {
                     hint:Query hint:optimizer 'None' .
                     {	BIND(entity:",entity_code," AS ?valUrl) .
                     BIND(\"N/A\" AS ?propUrl ) .
                     BIND(\"identity\"@en AS ?propLabel ) .
                     }
                     UNION
                     {	entity:", entity_code," ?propUrl ?valUrl .
                     ?property ?ref ?propUrl .
                     ?property rdf:type wikibase:Property .
                     ?property rdfs:label ?propLabel
                     }
                     
                     ?valUrl rdfs:label ?valLabel
                     FILTER (LANG(?valLabel) = 'en') .
                     OPTIONAL{ ?valUrl wdt:P18 ?picture .}
                     FILTER (lang(?propLabel) = 'en' )
                     }
                     ORDER BY ?propUrl ?valUrl
                     LIMIT 200")
    results <- quietly_query(query) 
    results <- results$result
    results$name<- name
    results
  }else{
    NULL
  }
   
  }


```

Yes, I just had to replace all occurrences of "sv" with "en" to get a function for this post. I'd like to try to write an automatic text about badass women. 

```{r}

get_a_string <- function(prop, prep, wikidata){
  answer <- dplyr::filter(wikidata, propLabel == prop) %>%
    .$valLabel %>%
    unique() %>%
    toString
  if(answer == ""){
    return("")
  }else{
    return(paste(prep, answer))
  }
}

tell_me_about <- function(name){
  wikidata <- get_wikidata(name)
  questions <- c("occupation", "country of citizenship",
                 "field of work", "award received")
  
  words <- c("a", "from", 
             "known from her work in", "and who was awarded")
  
  strings <- purrr::map2_chr(questions, words, 
                            get_a_string,
                            wikidata = wikidata)
  
  strings <- strings[strings != ""]
  
  sentence <- paste(name, "was", toString(strings))
  sentence <- paste0(sentence, ".")
  return(sentence)
}

```

Ok, let's try our automatic history writing function. It won't work for Dora and Joan, sadly.

```{r}
tell_me_about("Lyudmila Pavlichenko")
tell_me_about("Myra Hindley")
tell_me_about("Harriet Tubman")

```

Not many details clearly, but not too bad for a quickly written history hum bot, if I can call it so.

# So, happy, Nutella?

This was my contribution to the meme following Nutella's viral tweet. I am thankful for the badass women I did end up discovering thanks to the tweets, and am waiting for someone to replace the lyrics of all Taylor Swift's songs with gems from this Twitter trend.